---
title: "covid 19"
author: "Shauna Breazeal"
date: "3/18/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$: Covid 19 deaths in Korea Jan- May 2020 are unrelated to sex.
$H_A$: There is a relation between deaths in Korea Jan-May 2020 and sex of the patient.

Table of patients:
```{r}
kRdata$Sex = as.factor(kRdata$Sex)
kRdata$result= as.factor(kRdata$result)


# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent


```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=Sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The percentage barchart shows that about 35% of females die of Covid 19 in Korea while males are dying at a rate of about 65%, this would suggest that sex is a factor in dying due to Covid 19.



```{r}
barchartGC(~result + Sex,data=kRdata, main ="Frequency Bar Chart")
```

  
### this bar chart is displaying in total frequency terms our data.  Clearly there are a lot less dead people than alive people as is reflected by the sizes of the bars in the display.  





  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that 3 of every hundred males will die of Covid 19 in Korea, as opposed to 1.4 of every hundred females will die due to Covid 19

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


## Conclusions

The p-value is 0.001, the odds of a male dying due to Covid 19 in Korea are 2.3:1 compared to females, and the confidence interval does not contain a 1. Due to these findings we would reject the null hypothesis. 